---
title: "Honors Project Part 1"
author: "Amit Grewal"
date: "2/13/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r basic commands, echo=FALSE}
    #load "HOUSEDATA" dataset
library(readxl)
HOUSEDATA <- read_excel("C:/Users/Amit/OneDrive - Michigan State University/MKT 317/MKT 317 Data/HOUSEDATA.xlsx")
    #enable ggplot2 package
library(ggplot2)
    #enable scales package
library(scales)
```

# R Plots and Commands
## Plot 4
```{r plot 4}
ggplot(data=HOUSEDATA) +
  geom_point(mapping=aes(x=SIZE, y=PRICE), color="deepskyblue")
```

This plot graphs the datapoints from the "HOUSEDATA" dataset on the x~y planes SIZE~PRICE. SIZE corresponds to a datapoint/house's overall size in Sq. Ft., while PRICE corresponds to each house's list price in thousands of dollars.


## Plot 7
```{r plot 7}
ggplot(data=HOUSEDATA) +
  geom_point(mapping=aes(x=SIZE, y=PRICE, color=LOT)) +
  scale_color_manual(values=c("red", "deepskyblue"))
```

This plot graphs houses on the X~Y planes SIZE~PRICE which are the house's size in Sq. Ft. and its list price in thousands of dollars, respectively. The plot also differentiates among lot type with red dots corresponding to a corner lot and red dots to a non-corner lot.


## Plots 10-B, 10-C, and 10-D
Plot 10-A is not included as the R command had an error and thus did not compile properly.

### Plot 10-B
```{r plot 10-B, message=FALSE}
ggplot(data=HOUSEDATA) +
  geom_point(mapping=aes(x=SIZE, y=PRICE, color=LOT)) + 
  geom_smooth(mapping=aes(x=SIZE, y=PRICE), method="lm", se=FALSE)
```

This plot is incorrect as it only has one overall trend line instead of two trendlines, one for each lot type.

### Plot 10-C
```{r plot 10-C, message=FALSE}
ggplot(data=HOUSEDATA) +
  geom_point(mapping=aes(x=SIZE, y=PRICE)) + 
  geom_smooth(mapping=aes(x=SIZE, y=PRICE, color=LOT), method="lm", se=FALSE)
```

This plot is incorrect because the datapoints are not differentiated by lot type, and thus are not colored (other than the default black).

### Plot 10-D
```{r plot 10-D, message=FALSE}
ggplot(data=HOUSEDATA) +
  geom_point(mapping=aes(x=SIZE, y=PRICE, color=LOT)) + 
  geom_smooth(mapping=aes(x=SIZE, y=PRICE, color=LOT), method="lm", se=FALSE)
```

This plot is correct as it differentiates the datapoints/houses by lot type and includes a trend line for each the two groups of houses (corner and non-corner lots).

## Plot 17
```{r plot 17, message=FALSE}
ggplot(data=HOUSEDATA) +
  geom_point(mapping=aes(x=SIZE, y=PRICE, color=as.factor(AGENCY))) + 
  geom_smooth(mapping=aes(x=SIZE, y=PRICE, color=as.factor(AGENCY)), method="lm", se=FALSE) + 
  labs(x="Size (square feet)", y="Selling Price", color="Selling Agency") +
  scale_y_continuous(labels= label_dollar(prefix = "$", suffix = "K"))+
  theme_bw()
```

### Line 1
(ggplot...) plots the HOUSEDATA canvas.

### Line 2
(geom_point...) plots the datapoints/houses on the X~Y planes SIZE~PRICE (house size in Sq. Ft. and list price in thousands of dollars) and differentiates among them with color based on the selling agency.

### Line 3
(geom_smooth...) plots trend lines for each group of houses/datapoints (grouped according to selling agency) and colors them with the same color scheme as geom_point...

### Line 4
(labs..) changes the X-axis label to "Size (square feet)" and the Y-axis label to "Selling Price"

### Line 5
(scale_y...) changes the Y-axis units to "\$YYYK" where "YYY" is the house's list price. I.e. the command adds the prefix "\$" and the suffix "K" to denote dollars and thousand, respectively

### Line 6
(theme...) changes the plot background from white border/gray cells/white grid lines to black border/white cells/gray grid lines

## Plot 18
```{r plot 18, message=FALSE}
ggplot(data=HOUSEDATA) +
  geom_point(mapping=aes(x=SIZE, y=PRICE*1000, color=as.factor(AGENCY))) + 
  geom_smooth(mapping=aes(x=SIZE, y=PRICE*1000, color=as.factor(AGENCY)), method="lm", se=FALSE) + 
  labs(x="Size (square feet)", y="Selling Price", color="Selling Agency") + 
  scale_y_continuous(labels=scales::dollar_format()) + 
  theme_bw()
```







